### 2020-08-15 README.Rmd
add RepeatedSptCVCstf

---
output: github_document
bibliography: inst/references.bib
biblio-style: apalike
nocite: | 
  @brenning2012, @valavi2018, @schratz2019
---

# mlr3spatiotempcv

Spatio-temporal resampling methods for mlr3.

<!-- badges: start -->

[![tic](https://github.com/mlr-org/mlr3spatiotempcv/workflows/tic/badge.svg?branch=master)](https://github.com/mlr-org/mlr3spatiotempcv/actions) [![CRAN Status Badge](https://www.r-pkg.org/badges/version-ago/mlr3spatiotempcv)](https://cran.r-project.org/package=mlr3spatiotempcv) [![Coverage status](https://codecov.io/gh/mlr-org/mlr3spatiotempcv/branch/master/graph/badge.svg)](https://codecov.io/github/mlr-org/mlr3spatiotempcv?branch=master) [![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental) <!-- badges: end -->

This package extends the [mlr3](https://github.com/mlr-org/mlr3) package framework with spatiotemporal resampling and visualization methods.

⚠️⚠️⚠️\
The package is in very early stages and breaking changes withour further notice are expected. If you want to use if for your research, you might need to refactor your analysis along the way.\
⚠️⚠️⚠️

## Resampling methods

Currently, the following ones are implemented:

| Literature                           | Package    | Reference     | mlr3 Sugar                     |
|--------------------------------------|------------|---------------|--------------------------------|
| Spatial Buffering                    | blockCV    | Valavi 2019   | `rsmp("spcv-buffer")`          |
| Spatial Blocking                     | blockCV    | Valavi 2019   | `rsmp("spcv-block")`           |
| Spatial CV                           | sperrorest | Brenning 2012 | `rsmp("spcv-coords")`          |
| Environmental Blocking               | blockCV    | Valavi 2019   | `rsmp("spcv-env")`             |
| \-                                   | \-         | \-            | `rsmp("sptcv-cluto")`          |
| Leave-Location-and-Time-Out          | CAST       | Meyer 2018    | `rsmp("sptcv-cstf")`           |
| ---                                  | ---        | ---           | ---                            |
| Repeated Spatial Blocking            | blockCV    | Valavi 2019   | `rsmp("repeated-spcv-block")`  |
| Repeated Spatial CV                  | sperrorest | Brenning 2012 | `rsmp("repeated-spcv-coords")` |
| Repeated Env Blocking                | blockCV    | Valavi 2019   | `rsmp("repeated-spcv-env")`    |
| \-                                   | \-         | \-            | `rsmp("repeated-sptcv-cluto")` |
| Repeated Leave-Location-and-Time-Out | CAST       | Meyer 2018    | `rsmp("repeated-sptcv-cstf")`  |

## Spatial tasks

| Name     | Code              | Type    |
|----------|-------------------|---------|
| ecuador  | `tsk("ecuador")`  | Classif |
| diplodia | `tsk("diplodia")` | Classif |

## Spatiotemporal tasks

| Name     | Code              | Type |
|----------|-------------------|------|
| cookfarm | `tsk("cookfarm")` | Regr |

## Visualization

S3 `autoplot()` for all implemented spatial resampling methods.

### Visualization of all folds

```{r setup, include=FALSE}
set.seed(123)
knitr::opts_chunk$set(fig.path = "man/figures/README-")
```

```{r spcv-coords-all-partitions, warning=FALSE}
library(mlr3)
library(mlr3spatiotempcv)
library(ggplot2)

task = tsk("ecuador")
resampling = rsmp("spcv-coords", folds = 5)
resampling$instantiate(task)

autoplot(resampling, task)
```

### Visualization of a specific fold

```{r spcv-coords-fold}
autoplot(resampling, task, fold_id = 1)
```

### Spatiotemporal Visualization

Three-dimensional visualization via {plotly}

(See vignette ["Spatiotemporal Visualization"](https://mlr3spatiotempcv.mlr-org.com/articles/spatiotemp-viz.html) for an interactive 3D HTML variant.)

```{r spatiotemp, eval=FALSE}
library(plotly)
task_st = tsk("cookfarm")
resampling = rsmp("spcv-cluto", folds = 5)
resampling$instantiate(task_st, time_var = "Date")
autoplot(resampling, task_st, fold_id = 1, point_size = 3)
```

![](man/figures/spt-viz.png)

## More resources

For detailed information on how to use spatial resampling in {mlr3} please read the section about [spatial analysis in the mlr3 book](https://mlr3book.mlr-org.com/spatial.html) and consult the [Getting Started](https://mlr3spatiotempcv.mlr-org.com/articles/mlr3spatiotempcv.html) vignette.

# References